Outdated and License-Violating Code on Stack Overflow

### 2017-02-25 cloneoverflow-latex/cloneoverflow.tex
Back under 10 pages!

We call code snippets that are copied from software systems to online
Q\&A websites (such as Stack Overflow) and vice versa as ``online
code clones'' (sometimes only ``clones'' for brevity). There are two
directions in creating online code clones: (1) code is cloned from a software
project to a Q\&A website as an example; or (2) code is cloned from a
Q\&A website to a software project to obtain a functionality, perform
a particular task, or fixing a bug.
% and (3) code is transferred from one software project to another via a Q\&A website; 
Similar to classic clones, online code clones can lead to license
violations, bug propagation, introduction of vulnerabilities, and
re-use of outdated code. Unfortunately, online clones are more
difficult to locate and fix since the search space in online code
corpora is larger and no longer confined to a local repository.

@@

A motivating example of outdated online code clones can be found in an
answer to a Stack Overflow question regarding how to implement
{\small{\texttt{RawComparator}}} in
\textsf{hadoop}\footnote{\url{http://stackoverflow.com/questions/22262310}}.
\Cref{fig:before-after} shows a code snippet embedded as a part of the
accepted answer on the left. The snippet shows how \textsf{hadoop}
implements the {\small{\texttt{compare}}} method in its
{\small{\texttt{WritableComparator}}} class. The code snippet on the
right shows another version of the same method, but at this time
extracted from the latest version (as of 26 September 2016) of
\textsf{hadoop}. We can see that they are highly similar except a line
containing {\small{\verb|buffer.reset(null,0,0);|}} which was added in
the latest version. The added line is intended for cleaning up the
reference in the {\small{\verb|buffer|}} variable. While this change
has already been introduced into the {\small{\texttt{compare}}} method
in the latest version of \textsf{hadoop}, the code example in Stack
Overflow post is still unchanged. This example shows that there can be
inconsistencies between online code clones and their originals leading
users of Stack Overflow to reuse outdated code.
%The code snippet on Stack Overflow can be outdated. %Since reusing source code from Stack Overflow is considered a common practice nowadays, the scale of online code cloning is more than intra or inter project clone in a local code bases.
%This is an emerging and challenging problem. Since studies in this area are still limited, we aim to gain more insight of the problem in this study. %We are interested to gain more insights of online code clone in this study.

@@

%\subsection{Experimental Setup}
%\vspace{1ex}
To answer these four research questions, we perform an empirical study to
locate and study the online code clones between Stack Overflow and open source
projects. 
% We explain our process of building the data sets in each phase separately below.

%\subsection{Experimental Setup}

%We followed the experimental framework by processing the source code from the two data sets, configuring and running the clone detectors, and analysing the clones. We discuss each of them in details below.

\subsection{Methodology}

We designed our study in 6 phases as depicted in \Cref{fig:exp_framework} where
we build different data sets to answer each of our four research
questions.

\subsubsection{Phase 1: Clone Identification}

@@

\subsubsection{Phase 2: Clone Filtering}
As there are in total 315,786,118 clone pairs reported, it is
infeasible for humans to manually validate them all. Instead of
investigating a random sample, we adopted the idea of \textbf{clone
  agreement} which has been used in clone research
studies~\cite{Funaro2010, Wang2013,cr2016ssbse} in situations where a
clone oracle is missing or impossible to establish. Clone pairs agreed
by multiple clone detection tools have a higher likelihood to be real
clones~\cite{cr2016ssbse}. By using this agreement-based approach, we
reduced the number of clone candidates for manual investigation to the
ones agreed by multiple tools. To find agreement between two clone
pairs reported by two different tools, we used the clone pair matching
metric proposed by Bellon et al.~\cite{Bellon2007}. Two clone pairs
which have a large enough number of overlapping lines can be
categorised as either a good-match or an ok-match pair (denoted as
\textit{good} and \textit{ok} respectively) with a confidence value
between 0 and 1. A \textit{good} clone pair has stronger agreement
than an \textit{ok} pair. We follow the original definitions of good-
and ok-match introduced in Bellon's paper~\cite{Bellon2007}.

@@

%\subsubsection{Filtered Clones Set}
We selected all \textit{good} pairs (2,313) and the \textit{ok} pairs for
$S_D$--$N_D$ (12,336), $S_E$--$N_D$ (993), and $S_D$--$N_E$ (79), but we ignored
20,021 $S_EN_E$ \textit{ok} clone pairs in the classification due to
their large number of clones and the very small average clone sizes of
5.95 and 6.28 lines (see \Cref{tab:orig_stats}), leading to a high
probability of being trivial clones. We refer to the set of clone
pairs as the \textit{good} set and the \textit{ok'} set.
Together with the 17,801 (referred as $S_D'$ set) and 83 (referred as
$N_D'$ set) clone pairs after filtering, we ended up with a set of 32,533 online clone pair candidates.

@@

\subsubsection{Phase 3-4: Clone Validation and Classification}
We used the 3,636 filtered clone pairs after automatic classification for
manual validation and classification.
The validation and classification of the pairs was done at the same time. 
The clone validation process (phase 3 in \Cref{fig:exp_framework}) involves checking 
if a clone pair is a true positive or a false positive. 
Moreover, we are also interested in 
the patterns of code cloning so we can gain more insights into 
how these clones are created (phase 4 in \Cref{fig:exp_framework}).

@@

Regarding the online cloning classification patterns, 
we studied the 8 patterns of cloning from Kapser et
al.~\cite{Kapser2006,Kapser2008} and performed a preliminary study to
evaluate its applicability to our study. 
We tried to classify 697 online clone pairs from
the reported clones in phase 1 
%Using the $S_D$ clone report,
%snippets were ranked according to (1)~clone pair frequency,
%(2)~popularity (i.e.~number of associated Qualitas projects),
%(3)~clone size in SLOC, and (4)~ratio of cloned code per snippet. We
%selected these four criteria so that we could cover the online clones
%from various aspects. We selected the top 10 from each ranking and
%obtained 34 unique Stack Overflow snippets. The snippets were
%associated with 697 clone pairs.
%\subsubsection{Defining Online Code Cloning Patterns}
using Kapser's cloning patterns. We found that 
% the
% clone pairs were categorised into either Customisation or Templating.
% Thus, 
Kapser's patterns are too broad for our study and a
more suitable and fine-grained classification scheme is needed. After
the preliminary study, we adopted one of Kapser's cloning patterns,
\emph{boiler-plate code}, and defined
6 new cloning patterns. The seven patterns include QS, SQ, EX, UD, BP,
IN, and AC as presented in \Cref{tab:classification_scheme}. Pattern
QS (\textbf{Q}ualitas to \textbf{S}tack Overflow) represents clones
that have evidence of being copied from a Qualitas project to Stack
Overflow. The evidence of copying can be found in comments in the
Qualitas source code or in the Stack Overflow post's contents. Pattern
SQ (\textbf{S}tack Overflow to \textbf{Q}ualitas) is cloning, with
evidence, in the opposite direction from Stack Overflow to a Qualitas
project. Pattern EX (\textbf{Ex}ternal Sources) is cloning that has
evidence of copying from a single or multiple external sources to
Stack Overflow and to a Qualitas project.  Pattern UD
(\textbf{U}nknown \textbf{D}irection) is cloning that creates
identical or highly similar clones between Qualitas and Stack Overflow
but where we could not find any attribution of copying.
Pattern BP (\textbf{B}oiler-\textbf{P}late) represents clones
containing boiler-plate code of {\small\verb|equals()|} methods,
getters/setters, or IDE-generated code such as GUI components. Pattern
IN (\textbf{In}heritance/Interface) is cloning by inheritance of the
same super class or implementation of the same interface. These two
activities usually result in similar overriding methods. The last
pattern, AC (\textbf{A}ccidental \textbf{C}lones), represents
accidentally similar clone pairs. These are mainly false positive
clones from the clone detectors such as similar
{\small\texttt{try-catch}} statements.

@@

\subsubsection{Phase 5: Outdated Clones}
Outdated code occurs when a piece of code has been copied from its
origin to another location and later the original has been
updated~\cite{Xia2014}. Usually code clone detection is used to locate
clone instances and update them to match with the
originals~\cite{Bellon2007}. However, online code clones are more
difficult to detect than in regular software projects due to its large
search space and a mix of natural and programming languages combined
in the same post.
%Moreover, Stack Overflow is not a software. There is no test suite to run and check whether the validity of the code snippets. To the best of our knowledge, this work is the first to address the problem of outdated code on Stack Overflow.

@@

\subsubsection{Phase 6: Licensing Analysis}
Software licensing plays an important role in software
development. Violation of software licenses impacts software delivery
and also leads to legal issues~\cite{Sprigman2015}.
%It is an emerging area that software engineering research community is paying attention to. For example, 
One can run into a licensing issue if one integrates third-party
source code into their software without checking. A study by An et
al.~\cite{An2017} reports 1,219 cases of potential license violations
between 399 Android apps and Stack Overflow code
snippets.

@@

\textbf{EX: External Sources.} We found 200 online clone pairs with
evidence of cloning from external sources to Stack Overflow and to Qualitas.  
They include 4, 44, 146, and 6 pairs from the \textit{good}, \textit{ok'},
$S_D'$, and $N_D'$ set respectively. All of them contain statements
giving an external source of the cloned code.  For example, Stack
Overflow Post 9549009 contains a code comment saying \textit{``Copied
  shamelessly from
  org.bouncycastle.crypto.generators.PKCS5S2ParametersGenerator''}.

@@

\textbf{UD: Unknown Direction.} We found 359 online clone pairs with
no evidence of cloning between Qualitas and Stack Overflow but with a
high code similarity that suggests cloning. They include 15, 20, 293,
and 31 pairs from the \textit{good}, \textit{ok'}, $S_D'$, and $N_D'$
set respectively. The most cloned projects is \textsf{netbeans} with
93 clone pairs. Most of the clones are a large chunk of code handling
GUI components. Although these GUI clones might be auto-generated by
IDEs, we did not find any evidence. The second most cloned project is
\textsf{jung2} (62 pairs), a network and graph framework, followed by
\textsf{eclipse-SDK} (37 pairs).
% There are several Stack Overflow snippets that are identical to code
% in \textsf{jung2}.

@@

\begin{figure}
	\centering
	\includegraphics[width=\linewidth]{freshness}
	\caption{Status of the 112 QS online clone pairs}
	\label{fig:outdated}
\end{figure}

@@

%If developers copy and reuse these licensed pieces of code in their projects, conflicts may happen without their realisation.
In our study, we reveal another possible situation of software
licensing issues caused by code cloned to Stack Overflow. We found
evidence that 112 pieces of code have been copied from Qualitas
projects to Stack Overflow as examples. Their status of accepted
answers increase their chances of being reused. Even though most of
the Qualitas projects came with a software license, we found that the licensing information
were frequently missing after the code was copied to Stack
Overflow. The licensing terms on top of source code files are not
copied because usually only a small part of the file was
cloned. In overall, we can see that most of the Stack Overflow snippets do not
contain licensing terms while their clone counterparts in Qualitas
projects do. The summary of
licensing information is listed in \Cref{tab:license_abc}.

@@

\textbf{Incompatible license:} there are 358 clone pairs which do not
contain licensing information after they are posted on Stack Overflow
or contain a different license from their Qualitas clone
counterparts. More than half (79) of \textbf{QS} clone pairs have
their licensing terms removed or changed when posted on Stack
Overflow. For \textbf{EX} clone pairs, we searched for licensing terms of the
original source code from the external sources. We found that 77 out
of 88 EX clone pairs have incompatible licenses.
%They are clones that have been identified to be copied from an external source.
Similarly, the license statement was removed from Stack Overflow
snippets. Of 226 \textbf{UD} clone pairs, 202 pairs have incompatible
licenses. Again, most of the clones in Qualitas contain a license
while the Stack Overflow snippets do not.
%
%\Cref{fig:sankey_license} shows the direction of changes in online code clone license. The direction of the change is from left to right. The thickness of the line reflects the number of clones for that relationship. In \Cref{fig:sankey_license} (A), we can clearly see that most of the QS online clones have their licenses changed from having a license to ``No license'' on Stack Overflow. The same findings were found for EX online clones as depicted in \Cref{fig:sankey_license} (B). We found that most of the EX clones have their licensing information stripped of after copying to Stack Overflow. These clones are instead covered by the default CC BY-NC-SA 3.0 license of Stack Overflow which is strongly permissive. These used-to-be-licensed code are dangerous for being reused because they can implicitly conflict with the reused software's license.

>>>>

We call code snippets that are copied from software systems to online
Q\&A websites (such as Stack Overflow) and vice versa as ``online
code clones''. There are two
directions in creating online code clones: (1) code is cloned from a software
project to a Q\&A website as an example; or (2) code is cloned from a
Q\&A website to a software project to obtain a functionality, perform
a particular task, or fixing a bug.
% and (3) code is transferred from one software project to another via a Q\&A website; 
Similar to classic clones, online code clones can lead to license
violations, bug propagation, introduction of vulnerabilities, and
re-use of outdated code. Unfortunately, online clones are 
difficult to locate and fix since the search space in online code
corpora is larger and no longer confined to a local repository.

@@

A motivating example of outdated online code clones can be found in an
answer to a Stack Overflow question regarding how to implement
{\small{\texttt{RawComparator}}} in
\textsf{hadoop}\footnote{\url{http://stackoverflow.com/questions/22262310}}.
\Cref{fig:before-after} shows a code snippet embedded as a part of the
accepted answer on the left. The snippet shows how \textsf{hadoop}
implements the {\small{\texttt{compare}}} method in its
{\small{\texttt{WritableComparator}}} class. The code snippet on the
right shows another version of the same method, but at this time
extracted from the latest version (as of 26 September 2016) of
\textsf{hadoop}. We can see that they are highly similar except a line
containing {\small{\verb|buffer.reset(null,0,0);|}} which was added in
the latest version. The added line is intended for cleaning up the
reference in the {\small{\verb|buffer|}} variable. While this change
has already been introduced into the {\small{\texttt{compare}}} method
in the latest version of \textsf{hadoop}, the code example in Stack
Overflow post is still unchanged. This example shows that
inconsistencies between online code clones and their originals can
lead users of Stack Overflow to reuse outdated code.
%The code snippet on Stack Overflow can be outdated. %Since reusing source code from Stack Overflow is considered a common practice nowadays, the scale of online code cloning is more than intra or inter project clone in a local code bases.
%This is an emerging and challenging problem. Since studies in this area are still limited, we aim to gain more insight of the problem in this study. %We are interested to gain more insights of online code clone in this study.

@@

%\subsection{Experimental Setup}
%\vspace{1ex}
To answer these four research questions, we perform an empirical study to
locate and study the online code clones between Stack Overflow and open source
projects. 
%
%\subsection{Methodology}
%
We designed our study in 6 phases as depicted in \Cref{fig:exp_framework} where
we build different data sets to answer each of our four research
questions.

\subsection{Phase 1: Clone Identification}

@@

\subsection{Phase 2: Clone Filtering}
As there are in total 315,786,118 clone pairs reported, it is
infeasible for humans to manually validate them all. Instead of
investigating a random sample, we adopted the idea of \textbf{clone
  agreement} which has been used in clone research
studies~\cite{Funaro2010, Wang2013,cr2016ssbse} in situations where a
clone oracle is missing or impossible to establish. Clone pairs agreed
by multiple clone detection tools have a higher likelihood to be real
clones~\cite{cr2016ssbse}. By using this agreement-based approach, we
reduced the number of clone candidates for manual investigation to the
ones agreed by multiple tools. To find agreement between two clone
pairs reported by two different tools, we used the clone pair matching
metric proposed by Bellon et al.~\cite{Bellon2007}. Two clone pairs
which have a large enough number of overlapping lines can be
categorised as either a good-match or an ok-match pair (denoted as
\textit{good} and \textit{ok} respectively) with a confidence value
between 0 and 1. A \textit{good} clone pair has stronger agreement
than an \textit{ok} pair. We follow the original definitions of good-
and ok-match introduced in Bellon's paper~\cite{Bellon2007}.

@@

%\subsection{Filtered Clones Set}
We selected all \textit{good} pairs (2,313) and the \textit{ok} pairs for
$S_D$--$N_D$ (12,336), $S_E$--$N_D$ (993), and $S_D$--$N_E$ (79), but we ignored
20,021 $S_EN_E$ \textit{ok} clone pairs in the classification due to
their large number of clones and the very small average clone sizes of
5.95 and 6.28 lines (see \Cref{tab:orig_stats}), leading to a high
probability of being trivial clones. We refer to the set of clone
pairs as the \textit{good} set and the \textit{ok'} set.
Together with the 17,801 (referred as $S_D'$ set) and 83 (referred as
$N_D'$ set) clone pairs after filtering, we ended up with a set of 32,533 online clone pair candidates.

@@

\subsection{Phase 3-4: Validation and Classification}
We used the 3,636 filtered clone pairs after automatic classification for
manual validation and classification.
The validation and classification of the pairs was done at the same time. 
The clone validation process (phase 3 in \Cref{fig:exp_framework}) involves checking 
if a clone pair is a true positive or a false positive. 
Moreover, we are also interested in 
the patterns of code cloning so we can gain more insights into 
how these clones are created (phase 4 in \Cref{fig:exp_framework}).

@@

Regarding the online cloning classification patterns, 
we studied the 8 patterns of cloning from Kapser et
al.~\cite{Kapser2006,Kapser2008} and performed a preliminary study to
evaluate its applicability to our study. 
We tried to classify 697 online clone pairs from
the reported clones in phase 1 
%Using the $S_D$ clone report,
%snippets were ranked according to (1)~clone pair frequency,
%(2)~popularity (i.e.~number of associated Qualitas projects),
%(3)~clone size in SLOC, and (4)~ratio of cloned code per snippet. We
%selected these four criteria so that we could cover the online clones
%from various aspects. We selected the top 10 from each ranking and
%obtained 34 unique Stack Overflow snippets. The snippets were
%associated with 697 clone pairs.
%\subsection{Defining Online Code Cloning Patterns}
using Kapser's cloning patterns. We found that 
% the
% clone pairs were categorised into either Customisation or Templating.
% Thus, 
Kapser's patterns are too broad for our study and a
more suitable and fine-grained classification scheme is needed. After
the preliminary study, we adopted one of Kapser's cloning patterns,
\emph{boiler-plate code}, and defined
6 new cloning patterns. The seven patterns include QS, SQ, EX, UD, BP,
IN, and AC as presented in \Cref{tab:classification_scheme}. Pattern
QS (\textbf{Q}ualitas to \textbf{S}tack Overflow) represents clones
that have evidence of being copied from a Qualitas project to Stack
Overflow. The evidence of copying can be found in comments in the
Qualitas source code or in the Stack Overflow post's contents. Pattern
SQ (\textbf{S}tack Overflow to \textbf{Q}ualitas) is cloning, with
evidence, in the opposite direction from Stack Overflow to a Qualitas
project. Pattern EX (\textbf{Ex}ternal Sources) is cloning that has
evidence of copying from a single or multiple external sources to
Stack Overflow and to a Qualitas project.  Pattern UD
(\textbf{U}nknown \textbf{D}irection) is cloning that creates
identical or highly similar clones between Qualitas and Stack Overflow
but where we could not find any attribution of copying.
Pattern BP (\textbf{B}oiler-\textbf{P}late) represents clones
containing boiler-plate code of {\small\verb|equals()|} methods,
getters/setters, or IDE-generated code such as GUI components. Pattern
IN (\textbf{In}heritance/Interface) is cloning by inheritance of the
same super class or implementation of the same interface. These two
activities usually result in similar overriding methods. The last
pattern, AC (\textbf{A}ccidental \textbf{C}lones), represents
accidentally similar clone pairs. These are mainly false positive
clones from the clone detectors such as similar
{\small\texttt{try-catch}} statements.

@@

\subsection{Phase 5: Outdated Clones}
Outdated code occurs when a piece of code has been copied from its
origin to another location and later the original has been
updated~\cite{Xia2014}. Usually code clone detection is used to locate
clone instances and update them to match with the
originals~\cite{Bellon2007}. However, online code clones are more
difficult to detect than in regular software projects due to its large
search space and a mix of natural and programming languages combined
in the same post.
%Moreover, Stack Overflow is not a software. There is no test suite to run and check whether the validity of the code snippets. To the best of our knowledge, this work is the first to address the problem of outdated code on Stack Overflow.

@@

\subsection{Phase 6: Licensing Analysis}
Software licensing plays an important role in software
development. Violation of software licenses impacts software delivery
and also leads to legal issues~\cite{Sprigman2015}.
%It is an emerging area that software engineering research community is paying attention to. For example, 
One can run into a licensing issue if one integrates third-party
source code into their software without checking. A study by An et
al.~\cite{An2017} reports 1,219 cases of potential license violations
between 399 Android apps and Stack Overflow code
snippets.

@@

\textbf{EX: External Sources.} We found 200 clone pairs with
evidence of cloning from external sources to Stack Overflow and to Qualitas.  
They include 4, 44, 146, and 6 pairs from the \textit{good}, \textit{ok'},
$S_D'$, and $N_D'$ set respectively. All of them contain statements
giving an external source of the cloned code.  For example, Stack
Overflow Post 9549009 contains a code comment saying \textit{``Copied
  shamelessly from
  org.bouncycastle.crypto.generators.PKCS5S2ParametersGenerator''}.

@@

\textbf{UD: Unknown Direction.} We found 359 online clone pairs with
no evidence of cloning between Qualitas and Stack Overflow but with a
high code similarity that suggests cloning. 
% They include 15, 20, 293, and 31 pairs from the \textit{good},
% \textit{ok'}, $S_D'$, and $N_D'$ set respectively.
The most cloned projects is \textsf{netbeans} with
93 clone pairs. Most of the clones are a large chunk of code handling
GUI components. Although these GUI clones might be auto-generated by
IDEs, we did not find any evidence. The second most cloned project is
\textsf{jung2} (62 pairs), a network and graph framework, followed by
\textsf{eclipse-SDK} (37 pairs).
% There are several Stack Overflow snippets that are identical to code
% in \textsf{jung2}.

@@

\begin{figure}
	\centering
	\includegraphics[width=\linewidth]{freshness}\vspace{-1ex}
	\caption{Status of the 112 QS online clone pairs}
	\label{fig:outdated}
\end{figure}

@@

%If developers copy and reuse these licensed pieces of code in their projects, conflicts may happen without their realisation.
In our study, we reveal another possible situation of software
licensing issues caused by code cloned to Stack Overflow. We found
evidence that 112 pieces of code have been copied from Qualitas
projects to Stack Overflow as examples. Their status of accepted
answers increase their chances of being reused. Even though most of
the Qualitas projects came with a software license, we found that the license information
were frequently missing after the code was copied to Stack
Overflow. The licensing terms on top of source code files are not
copied because usually only a small part of the file was
cloned. In overall, we can see that most of the Stack Overflow snippets do not
contain licensing terms while their clone counterparts in Qualitas
projects do. The summary of
licensing information is listed in \Cref{tab:license_abc}.

@@

\textbf{Incompatible license:} there are 358 clone pairs which do not
contain licensing information after they are posted on Stack Overflow
or contain a different license from their Qualitas clone
counterparts. More than half (79) of \textbf{QS} clone pairs have
their licensing terms removed or changed when posted on Stack
Overflow. For \textbf{EX} clone pairs, we searched for licensing terms of the
original source code from the external sources. We found that 77 out
of 88 EX clone pairs have incompatible licenses.
%They are clones that have been identified to be copied from an external source.
Similarly, the license statement was removed from Stack Overflow
snippets. Of 226 \textbf{UD} clone pairs, 202 pairs have incompatible
licenses. Again, most clones in Qualitas contain a license
while the Stack Overflow snippets do not.
%
%\Cref{fig:sankey_license} shows the direction of changes in online code clone license. The direction of the change is from left to right. The thickness of the line reflects the number of clones for that relationship. In \Cref{fig:sankey_license} (A), we can clearly see that most of the QS online clones have their licenses changed from having a license to ``No license'' on Stack Overflow. The same findings were found for EX online clones as depicted in \Cref{fig:sankey_license} (B). We found that most of the EX clones have their licensing information stripped of after copying to Stack Overflow. These clones are instead covered by the default CC BY-NC-SA 3.0 license of Stack Overflow which is strongly permissive. These used-to-be-licensed code are dangerous for being reused because they can implicitly conflict with the reused software's license.
